---
title: "clean_metadata"
author: "Payel Sardar"
format: pdf
editor: visual
---

### Importing necessary libraries

```{r}
library(stringr)
library(tidyr)
```

### Setting the working directory

```{r}
setwd("/users/k24034333/DCDM_Group6/")
```

### Setting the path to the metadata directory

```{r}
metadata_dir = "/users/k24034333/DCDM_Group6/metadata/"
```

### Reading and cleaning the files

1.  IMPC_procedure.txt

    Reading the lines of the file

    ```{r}
    proc_lines <- readLines(paste(metadata_dir,"IMPC_procedure.txt",sep=""))
    ```

    ```{r}
    data <- lapply(proc_lines[-1], function(line) {
      
      # Match the first field (line_number + procedureId) and the rest
      matches <- str_match(
        line,
        '^"([^"]+)"\\s+"([^,]+),\\s*(.*),\\s*(TRUE|FALSE),\\s*(\\d+)"$'
        )
      
      # Return the extracted fields
      c(matches[2], matches[3], matches[4], matches[5], matches[6])
    })
    ```

    Combining into a data frame

    ```{r}
    procedure_df <- do.call(rbind, data)
    colnames(procedure_df) <- c("line_number_procedureId",
                                "name",
                                "description",
                                "isMandatory",
                                "impcParameterOrigId")
    ```

    Saving the cleaned data in a CSV file

    ```{r}
    write.csv(procedure_df,
              paste(metadata_dir,"IMPC_procedure.csv",
                    sep=""),
              row.names = FALSE)
    ```

2.  IMPC_parameter_description.txt

    Reading the lines of the file

    ```{r}
    param_lines <- readLines(paste(metadata_dir,
                                   "IMPC_parameter_description.txt",
                                   sep=""))
    ```

    Using regex for separating data into columns

    ```{r}
    data <- lapply(param_lines[-1], function(line) {
      
      # Matching the first field and the rest
      matches <- str_match(
        line,
        '^"([^"]+)"\\s+"([^,]+),\\s*([^,]+),\\s*([^,]+),\\s*([^"]+)"$'
        )
      
      # Returning the extracted fields
      c(matches[2], matches[3], matches[4], matches[5], matches[6])
    })
    ```

    Combining into a dataframe

    ```{r}
    param_df <- do.call(rbind, data)

    # Using the first row as col names
    colnames(param_df) <- param_df[1, ]

    # Dropping the first row
    param_df <- param_df[-1, ]

    # Viewing the resulting dataframe
    head(param_df)
    ```

    Saving the resulting data frame in a CSV file

    ```{r}
    write.csv(param_df,
              paste(metadata_dir,"IMPC_parameter_description.csv",
                    sep=""),
              row.names = FALSE)
    ```
